---
title: "Short software tutorial on `nonprobsvy` package"
author: Maciej Beręsewicz
format: 
  html:
    number-sections: true
    self-contained: true
    toc: true
lang: en
bibliography: references.bib
biblio-style: apalike
execute: 
  warning: false
---

# Basic information

This tutorial shows basic usage of the
[nonprobsvy](https://github.com/ncn-foreigners/nonprobsvy) package
developed in this project. This package implements mass imputation,
inverse probability weighting and doubly robust estimators based on the
following papers:

-   Yang, S., Kim, J. K., & Song, R. (2020). Doubly robust inference
    when combining probability and non-probability samples with high
    dimensional data. Journal of the Royal Statistical Society. Series
    B, Statistical Methodology, 82(2), 445.
-   Kim, J. K., Park, S., Chen, Y., & Wu, C. (2021). Combining
    non-probability and probability survey samples through mass
    imputation. Journal of the Royal Statistical Society Series A:
    Statistics in Society, 184(3), 941-963.
-   Chen, Y., Li, P., & Wu, C. (2020). Doubly robust inference with
    nonprobability survey samples. Journal of the American Statistical
    Association, 115(532), 2011-2021.

Tutorial was prepared based on the development version of the
`nonprobsvy` (31.05.2023, commit
`748b20fc6c18b2889fce626a9742824d6fc1a34b`).

# Install and load the required packages

Install `remotes` package and then install `nonprobsvy` package from
github repository.

```{r install-packages, eval=FALSE}
install.packages("remotes")
remotes::install_github("ncn-foreigners/nonprobsvy")
install.packages("survey")
```

Load required packages

```{r load-packages}
library(survey)
library(nonprobsvy)
```

In the empirical part we reproduce results from the following papers:
@yang2020; @kim2021; @chen2020.

# Empirical examples 

## @yang2020 paper 

### Simulation from section 6

In this section, we evaluate the finite sample performance of the procedure proposed. We first
generate a finite population $\mathcal{F}_{N}=\left\{(X_{i}, Y_{i}): i=1, \ldots, N\right\}$ with $N=10000$, where $Y_{i}$ is a continuous or binary outcome variable, and $X_{i}=(1, X_{1, i}, \ldots, X_{p-1, i})^{\mathrm{T}}$ is a $p$-dimensional vector of covariates with the first component being 1 and other components independently generated from the standard normal distribution. We set $p=50$. From the finite population, we select a non-probability sample $\mathcal{B}$ of size $n_{\mathrm{B}} \approx 2000$, according to the selection indicator $I_{\mathrm{B}, i} \sim \operatorname{Ber}(\pi_{\mathrm{B}, i})$.

We select a probability sample $\mathcal{A}$ of the average size $n_{\mathrm{A}}=500$ under Poisson sampling with $\pi_{\mathrm{A}, i} \propto(0.25+\left|X_{1 i}\right|+0.03\left|Y_{i}\right|)$. The parameter of interest is the population mean $\mu=N^{-1} \sum_{i=1}^{N} Y_{i}$.

For the non-probability sampling probability, we consider both linear and non-linear sam-
pling score models:

+ (a) $\operatorname{logit}(\pi_{\mathrm{B}, i})=\alpha_{0}^{\mathrm{T}} X_{i}$, where $\alpha_{0}=(-2,1,1,1,1,0,0,0, \ldots, 0)^{\mathrm{T}}($ model PSM I)
+ (b) $\operatorname{logit}(\pi_{\mathrm{B}, i})=3.5+\alpha_{0}^{\mathrm{T}} \log (X_{i}^{2})-\sin (X_{3, i}+X_{4, i})-X_{5, i}-X_{6, i}$, where $\alpha_{0}=(0,0,0,3,3,3,3$
$0, \ldots, 0)^{\mathrm{T}}($ model PSM II).

For generating a continuous outcome variable $Y_{i}$, we consider both linear and non-linear
outcome models with $\beta_{0}=(1,0,0,1,1,1,1,0, \ldots, 0)^{\mathrm{T}}$ :

+ (a) $Y_{i}=\beta_{0}^{\mathrm{T}} X_{i}+\epsilon_{i}, \epsilon_{i} \sim \mathcal{N}(0,1)($ model OM I);
+ (b) $Y_{i}=1+\exp \left\{3 \sin (\beta_{0}^{\mathrm{T}} X_{i})\right\}+X_{5, i}+X_{6, i}+\epsilon_{i}, \epsilon_{i} \sim \mathcal{N}(0,1)$ (model OM II).

For generating a binary outcome variable $Y_{i}$, we consider both linear and non-linear outcome
models with $\beta_{0}=(1,0,0,3,3,3,3,0, \ldots, 0)^{\mathrm{T}}$,

+ (a) $Y \sim \operatorname{Ber}\left\{\pi_{Y}(X)\right\}$ with logit $\left\{\pi_{Y}(X)\right\}=\beta_{0}^{\mathrm{T}} X$ (model OM III);
+ (b) $Y \sim \operatorname{Ber}\left\{\pi_{Y}(X)\right\}$ with logit $\left\{\pi_{Y}(X)\right\}=2-\log \left\{(\beta_{0}^{\mathrm{T}} X)^{2}\right\}+2 X_{5, i}+2 X_{6, i}$ (model OM IV).

We consider the following estimators

+ (a) naive, $\hat{\mu}_{\text {naive }}$, the naive estimator using the simple average of $Y_{i}$ from sample $\mathrm{B}$, which
provides the degree of the selection bias of sample $\mathrm{B}$;
+ (b) oracle, $\hat{\mu}_{\text {ora }}$, the doubly robust estimator $\hat{\mu}_{\mathrm{dr}}(\hat{\alpha}_{\text {ora }}, \hat{\beta}_{\text {ora }})$, where $\hat{\alpha}_{\text {ora }}$ and $\hat{\beta}_{\text {ora }}$ are based
on the joint estimator restricting to the known important covariates for comparison;
+ (c) $p-ipw$, $\hat{\mu}_{\mathrm{p} \text {-ipw }}$, the penalized inverse probability of sampling weighting estimator $\hat{\mu}_{\mathrm{IPW}}=$ $N^{-1} \Sigma_{i \in \mathcal{B}} \hat{\pi}_{\mathrm{B}, i}^{-1} Y_{i}$, where logit $(\hat{\pi}_{\mathrm{B}, i})=X_{i}^{\mathrm{T}} \hat{\alpha}$ using a logistic regression model, and $\hat{\alpha}$ is obtained by a weighted penalized regression of $I_{\mathrm{B}, i}$ on $X_{i}$ based on the combined data from sample $\mathrm{A}$ and sample $\mathrm{B}$, with the units in sample A weighted by the known sampling weights and the units in sample $\mathrm{B}$ weighted by 1 .
+ (d) $p-reg$, $\hat{\mu}_{\mathrm{p} \text {-reg }}$, the penalized regression estimator $\hat{\mu}_{\mathrm{p}-\mathrm{reg}}=N^{-1} \Sigma_{i \in \mathcal{A}} d_{\mathrm{A}, i} m(X ; \hat{\beta})$, where $\hat{\beta}$ is obtained by a penalized regression of $Y_{i}$ on $X_{i}$ based on sample $\mathrm{B}$;
+ (e) $p-dr0$, $\hat{\mu}_{\mathrm{p} \text {-dr } 0}$, the penalized double estimating equation estimator based on the set of outcome predictors $\hat{\mathcal{M}}_{\beta}$;
+ (f) $p-dr$, $\hat{\mu}_{\mathrm{p}-\mathrm{dr}}$, the proposed penalized double estimating equation estimator based on the
union of sampling and outcome predictors $\hat{\mathcal{M}}_{\alpha} \cup \hat{\mathcal{M}}_{\beta}$.

We also note that $\hat{\mu}_{\mathrm{dr}}$ without variable selection is severely biased and unstable and therefore is excluded for comparison.


## @chen2020 paper 

## @kim2021 paper

### Simulation from section 6 

The setup for the simulation study employed a $3 \times 2$ factorial structure with two factors. The first factor is the superpopulation model that generates the finite population. The second factor is the
sample size for sample B. We used two levels for $n_{B}$, where $n_{B}=500$ or $n_{B}=1000$. We generated the following three models for finite populations of size $N=100,000$. The variables, $x_{i}$ and $e_{i}$, are independently generated from $N(2,1)$ and $N(0,1)$, respectively.

The study variable $y_{i}$ are constructed differently for each model:

+ Model I -- $y_{i}=1+2 x_{i}+e_{i}$
+ Model II -- $y_{i}=3+x_{i}+2e_{i}$
+ Model III -- $y_{i}=2.5 + 0.5x_{i}^2+e_{i}$

Model I generates a finite population with a high correlation between $x$ and $y$ ($r^{2}=0.8$), Model II generates a finite population with a low correlation ($r^{2}=0.2$), and Model III generates a finite population where the linear relationship fails. Model III is included to check the effect of model mis-specification for the imputation model.

From each of the three populations, we generated two independent samples. We use simple random
sampling of size $n_{A}=500$ to obtain sample $A$. In selecting sample $B$ of size $n_{B}$, where $n_{B} \in\{500,1000\}$,
we create two strata where Stratum 1 consists of elements with $x_{i} \leq 2$ and Stratum 2 consists of
elements with $x_{i}>2$. Within each stratum, we select $n_{h}$ elements by simple random sampling, in-
dependent between the two strata, where $n_{1}=0.7 n_{B}$ and $n_{2}=0.3 n_{B}$. We assume that the stratum
information is unavailable at the time of data analysis. Using the two samples $A$ and $B$, we compute
four estimators of $\theta_{N}=N^{-1} \sum_{i=1}^{N} y_{i}$ :

1. The sample mean from sample $A$ : $\widehat{\theta}_{A}=n_{A}^{-1} \sum_{i \in A} y_{i}$.
2. The naive estimator (sample mean) from sample $B: \widehat{\theta}_{B}=n_{B}^{-1} \sum_{i \in B} y_{i}$.
3. The mass imputation estimator from sample $A$ given in Equation (6) using $\widehat{y}_{i}=\widehat{\beta}_{0}+\widehat{\beta}_{1} x_{i}$ where
$(\widehat{\beta}_{0}, \widehat{\boldsymbol{\beta}}_{1})$ are the estimated regression coefficients obtained from sample $B$.
4. The IPW estimator proposed by @chen2020: $\widehat{\theta}_{I P W}=N^{-1} \sum_{i \in B} \widehat{\pi}_{i}^{-1} y_{i}$, where the propensity
scores, $\pi_{i}=\pi(\mathbf{x}_{i} ; \boldsymbol{\phi})=\left\{1+\exp (-\phi_{0}-\phi_{1} x_{i})\right\}^{-1}$ with $\mathbf{x}_{i}=(1, x_{i})^{\prime}$ and $\boldsymbol{\phi}=(\phi_{0}, \phi_{1})^{\prime}$, are estimated
by using $\widehat{\boldsymbol{\phi}}$ which solves the following score equations:
$$
U(\boldsymbol{\phi})=\sum_{i \in B} \mathbf{x}_{i}-\sum_{i \in A} w_{i} \pi(\mathbf{x}_{i} ; \boldsymbol{\phi}) \mathbf{x}_{i}=\mathbf{0}
$$


# Literature
