---
title: "Short software tutorial on `nonprobsvy` package"
author: Maciej Beręsewicz
format: html
lang: en
bibliography: references.bib
biblio-style: apalike
execute: 
  warning: false
---

# Basic information

This tutorial shows basic usage of the
[nonprobsvy](https://github.com/ncn-foreigners/nonprobsvy) package
developed in this project. This package implements mass imputation,
inverse probability weighting and doubly robust estimators based on the
following papers:

-   Yang, S., Kim, J. K., & Song, R. (2020). Doubly robust inference
    when combining probability and non-probability samples with high
    dimensional data. Journal of the Royal Statistical Society. Series
    B, Statistical Methodology, 82(2), 445.
-   Kim, J. K., Park, S., Chen, Y., & Wu, C. (2021). Combining
    non-probability and probability survey samples through mass
    imputation. Journal of the Royal Statistical Society Series A:
    Statistics in Society, 184(3), 941-963.
-   Chen, Y., Li, P., & Wu, C. (2020). Doubly robust inference with
    nonprobability survey samples. Journal of the American Statistical
    Association, 115(532), 2011-2021.

Tutorial was prepared based on the development version of the
`nonprobsvy` (31.05.2023, commit
`748b20fc6c18b2889fce626a9742824d6fc1a34b`).

# Install and load the required packages

Install `remotes` package and then install `nonprobsvy` package from
github repository.

```{r install-packages, eval=FALSE}
install.packages("remotes")
remotes::install_github("ncn-foreigners/nonprobsvy")
install.packages("survey")
```

Load required packages

```{r load-packages}
library(survey)
library(nonprobsvy)
```

In the empirical part we reproduce results from the following papers:
@yang2020; @kim2021; @chen2020.

# Empirical examples 

## @yang2020 paper 



## @chen2020 paper 

## @kim2021 paper

### Simulation from section 6 

The setup for the simulation study employed a $3 \times 2$ factorial structure with two factors. The first factor is the superpopulation model that generates the finite population. The second factor is the
sample size for sample B. We used two levels for $n_{B}$, where $n_{B}=500$ or $n_{B}=1000$. We generated the following three models for finite populations of size $N=100,000$. The variables, $x_{i}$ and $e_{i}$, are independently generated from $N(2,1)$ and $N(0,1)$, respectively.

The study variable $y_{i}$ are constructed differently for each model:

+ Model I -- $y_{i}=1+2 x_{i}+e_{i}$
+ Model II -- $y_{i}=3+x_{i}+2e_{i}$
+ Model III -- $y_{i}=2.5 + 0.5x_{i}^2+e_{i}$

Model I generates a finite population with a high correlation between $x$ and $y$ ($r^{2}=0.8$), Model II generates a finite population with a low correlation ($r^{2}=0.2$), and Model III generates a finite population where the linear relationship fails. Model III is included to check the effect of model mis-specification for the imputation model.

From each of the three populations, we generated two independent samples. We use simple random
sampling of size $n_{A}=500$ to obtain sample $A$. In selecting sample $B$ of size $n_{B}$, where $n_{B} \in\{500,1000\}$,
we create two strata where Stratum 1 consists of elements with $x_{i} \leq 2$ and Stratum 2 consists of
elements with $x_{i}>2$. Within each stratum, we select $n_{h}$ elements by simple random sampling, in-
dependent between the two strata, where $n_{1}=0.7 n_{B}$ and $n_{2}=0.3 n_{B}$. We assume that the stratum
information is unavailable at the time of data analysis. Using the two samples $A$ and $B$, we compute
four estimators of $\theta_{N}=N^{-1} \sum_{i=1}^{N} y_{i}$ :

1. The sample mean from sample $A$ : $\widehat{\theta}_{A}=n_{A}^{-1} \sum_{i \in A} y_{i}$.
2. The naive estimator (sample mean) from sample $B: \widehat{\theta}_{B}=n_{B}^{-1} \sum_{i \in B} y_{i}$.
3. The mass imputation estimator from sample $A$ given in Equation (6) using $\widehat{y}_{i}=\widehat{\beta}_{0}+\widehat{\beta}_{1} x_{i}$ where
$\left(\widehat{\beta}_{0}, \widehat{\boldsymbol{\beta}}_{1}\right)$ are the estimated regression coefficients obtained from sample $B$.
4. The IPW estimator proposed by @chen2020: $\widehat{\theta}_{I P W}=N^{-1} \sum_{i \in B} \widehat{\pi}_{i}^{-1} y_{i}$, where the propensity
scores, $\pi_{i}=\pi\left(\mathbf{x}_{i} ; \boldsymbol{\phi}\right)=\left\{1+\exp \left(-\phi_{0}-\phi_{1} x_{i}\right)\right\}^{-1}$ with $\mathbf{x}_{i}=\left(1, x_{i}\right)^{\prime}$ and $\boldsymbol{\phi}=\left(\phi_{0}, \phi_{1}\right)^{\prime}$, are estimated
by using $\widehat{\boldsymbol{\phi}}$ which solves the following score equations:
$$
U(\boldsymbol{\phi})=\sum_{i \in B} \mathbf{x}_{i}-\sum_{i \in A} w_{i} \pi\left(\mathbf{x}_{i} ; \boldsymbol{\phi}\right) \mathbf{x}_{i}=\mathbf{0}
$$


# Literature
